Probabilistic Catalogs for Crowded Stellar Fields

### 2012-10-08 paper/ms.tex
Tidying of figure placement and size
A few minor tweaks due to Hogg comments

\documentclass[letterpaper, 11pt]{article}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citet and citep
\usepackage{amsmath}	% for \iint
\usepackage{bbm}	% for blackboard bold numbers
\usepackage[left=3cm,top=3cm,right=3cm]{geometry}

@@

\title{Probabilistic Catalogs from Crowded Star Fields}
\author{Brendon J. Brewer$^{1,*}$, David W. Hogg$^{2}$,
and Daniel Foreman-Mackey$^{2}$ \\
\\
\small
$^1$ Department of Statistics, The University of Auckland,\\
\small
Private Bag 92019, Auckland 1142, New Zealand \\
\small
$^2$ Center for Cosmology and Particle Physics, Department of Physics,
New York University,\\
\small
4 Washington Place, New York, NY, 10003, USA\\
\small
$^*$\texttt{bj.brewer@auckland.ac.nz}
}

@@

\section{Introduction}
A fundamental problem in astronomy is the construction of {\it catalogs} from
raw image data. {\bf The catalogs are then used for...}
Standard tools for generating catalogs include \texttt{SExtractor}
\citep{sextractor}, {\bf more refs}.

@@

\subsection{The Prior Distribution}
The prior distribution for the number of stars $N$ is assigned to be uniform between 0 and
some maximum number $N_{\rm max}$. The extent of the image is assumed to be
from $x=-1$ to $x=1$ and from $y=-1$ to $y=1$ in arbitrary units, and the
positions of the stars are assigned independent uniform priors:
\begin{eqnarray}
x_i &\sim& \textnormal{Uniform}(-1.1, 1.1) \\
y_i &\sim& \textnormal{Uniform}(-1.1, 1.1) \\
\end{eqnarray}
The model for the luminosity function is a broken power-law with four
free parameters:
\begin{eqnarray}
\beta = \{h_1, h_2, \alpha_1, \alpha_2\}.
\end{eqnarray}
where $h_1$ is the lower flux cutoff, $h_2$ is a break-point, $\alpha_1$ is
the slope of the distribution between $h_1$ and $h_2$, and $\alpha_2$ is the
slope of the distribution above $h_2$. For details on the broken power-law
model, see Appendix~\ref{power_law}. The prior distribution on $h_1$, $h_2$,
$\alpha_1$ and $\alpha_2$ is assigned to be:
\begin{eqnarray}
\log h_1 &\sim& \textnormal{Uniform}(\log(10^{-3}), \log(10^{+3})) \\
\log h_2 &\sim& \textnormal{Uniform}(h_1, h_1 + 2.3) \\
\alpha_1 &\sim& \textnormal{Uniform}(1, 5) \\
\alpha_2 &\sim& \textnormal{Uniform}(1, 5).
\end{eqnarray}

@@

\begin{figure}
\hspace{-1cm}
\includegraphics{Figures/test_cases.eps}
\caption{Two simulated images.
{\bf Left}: A field containing $\sim$100 stars.
{\bf Right}: A field containing $\sim$1000 stars.\label{fig:simulated_data}}
\end{figure}

@@

\section{Broken Power-Law Distribution}\label{power_law}
The broken power-law distribution is based on a straightforward extension
to a simple power-law distribution (also known as a Pareto distribution,
particularly in the statistics literature). The power-law distribution for a
variable $x$ (given a lower cutoff $x=h$ and a slope $\alpha$) is defined by:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcr}
0, & & x < h \\
x^{-\alpha - 1} & & x \geq h.
\end{array}
\right.
\end{eqnarray}
In contrast, the broken power-law distribution for a variable $x$ is defined by
a lower cutoff $x=h_1$, two slopes $\{\alpha_1, \alpha_2\}$ and a break point
$x=h_2$:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
x^{-\alpha_1 - 1} & & h_1 \leq x \leq h_2 \\
x^{-\alpha_2 - 1} & & x > h_2.
\end{array}
\right.
\end{eqnarray}
The free parameters of the broken power-law are:
\begin{eqnarray}
\beta = \{h_1, h_2, \alpha_1, \alpha_2\}.
\end{eqnarray}
With normalising terms included, the proportionality becomes an equality:
\begin{eqnarray}
p(x) &=&
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
Z_1^{-1}x^{-\alpha_1 - 1}, & & h_1 \leq x \leq h_2 \\
Z_2^{-1}x^{-\alpha_2 - 1}, & & x > h_2.
\end{array}
\right.
\end{eqnarray}
Two conditions will be used to determine the normalisers $Z_1$ and $Z_2$.
Firstly, the PDF should be continuous at $x=h_2$:
\begin{eqnarray}
Z_1^{-1}h_2^{-\alpha_1 - 1} &=& Z_2^{-1}h_2^{-\alpha_2 - 1}\\
\implies
Z_2 &=& Z_1h_2^{\alpha_1-\alpha_2}
\end{eqnarray}
The second condition is that the total probability must be 1:
\begin{eqnarray}
\int_{h_1}^{h_2} Z_1^{-1} x^{-\alpha_1 - 1} \, dx
+
\int_{h_2}^\infty Z_2^{-1} x^{-\alpha_2 - 1} \, dx
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
Z_2^{-1}\alpha_2^{-1}h_2^{-\alpha_2}
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
Z_1^{-1}h_2^{\alpha_2-\alpha_1}\alpha_2^{-1}h_2^{-\alpha_2}
&=& 1
\end{eqnarray}
\begin{eqnarray}
\implies
Z_1 &=& \alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
h_2^{-\alpha_1}\alpha_2^{-1}.
\end{eqnarray}
The cumulative distribution (CDF) is:
\begin{eqnarray}
P(X \leq x) = F(x) &=& 
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
(Z_1\alpha_1)^{-1}\left(h_1^{-\alpha_1} - x^{-\alpha_1}\right), & & h_1 \leq x \leq h_2 \\
1 - (Z_2\alpha_2)^{-1}x^{-\alpha_2}, & & x > h_2 
\end{array}
\right.
\end{eqnarray}
and the inverse CDF is:
\begin{eqnarray}
F^{-1}(u) &=& 
\left\{
\begin{array}{lcr}
\left[h_1^{-\alpha_1} - uZ_1\alpha_1\right]^{-1/\alpha_1}, & & 0 < u < 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}\\
\left[Z_2\alpha_2(1-u)\right]^{-1/\alpha_2},& & 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}< u < 1
\end{array}
\right.
\end{eqnarray}
An example of a broken power-law distribution is shown in Figure~\ref{fig:powerlaw}.
\begin{figure}[h!]
\begin{center}
\includegraphics[scale=0.7]{Figures/broken.eps}
\caption{A broken power-law distribution.\label{fig:powerlaw}}
\end{center}
\end{figure}

@@

\subsubsection{Proposal Distributions}

@@

>>>>

\documentclass[letterpaper, 11pt]{article}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citet and citep
\usepackage{amsmath}	% for \iint
\usepackage{bbm}	% for blackboard bold numbers
\usepackage[left=2cm,top=3cm,right=2cm]{geometry}

@@

\title{Probabilistic Catalogs for Crowded Stellar Fields}
\author{Brendon J. Brewer$^{1,*}$, David W. Hogg$^{2}$,
and Daniel Foreman-Mackey$^{2}$ \\
\\
\small
$^1$ Department of Statistics, The University of Auckland,\\
\small
Private Bag 92019, Auckland 1142, New Zealand \\
\small
$^2$ Center for Cosmology and Particle Physics, Department of Physics,
New York University,\\
\small
4 Washington Place, New York, NY, 10003, USA\\
\small
$^*$\texttt{bj.brewer@auckland.ac.nz}
}

@@

\section{Introduction}
A fundamental problem in astronomy is the construction of {\it catalogs} from
raw image data. {\bf The catalogs are then used for...}
Standard tools for generating catalogs include \texttt{SExtractor}
\citep{sextractor}, {\bf more refs}.
\citep{2011EAS....45..351H}

@@

\subsection{The Prior Distribution}
The prior distribution for the number of stars $N$ is assigned to be uniform between 0 and
some maximum number $N_{\rm max}$. The extent of the image is assumed to be
from $x=-1$ to $x=1$ and from $y=-1$ to $y=1$ in arbitrary units, and the
positions of the stars are assigned independent uniform priors:
\begin{eqnarray}
x_i &\sim& \textnormal{Uniform}(-1.1, 1.1) \\
y_i &\sim& \textnormal{Uniform}(-1.1, 1.1)
\end{eqnarray}
The stars are allowed to be slightly outside of the observed image because the
PSF can scatter light from these stars into the image. The model for the luminosity function is a broken power-law with four
free parameters:
\begin{eqnarray}
\beta = \{h_1, h_2, \alpha_1, \alpha_2\}.
\end{eqnarray}
where $h_1$ is the lower flux limit, $h_2$ is a break-point, $\alpha_1$ is
the slope of the distribution between $h_1$ and $h_2$, and $\alpha_2$ is the
slope of the distribution above $h_2$. For details on the broken power-law
model, see Appendix~\ref{power_law}. The prior distribution on $h_1$, $h_2$,
$\alpha_1$ and $\alpha_2$ is assigned to be:
\begin{eqnarray}
\log h_1 &\sim& \textnormal{Uniform}(\log(10^{-3}), \log(10^{+3})) \\
\log h_2 &\sim& \textnormal{Uniform}(h_1, h_1 + 2.3) \\
\alpha_1 &\sim& \textnormal{Uniform}(1, 5) \\
\alpha_2 &\sim& \textnormal{Uniform}(1, 5).
\end{eqnarray}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{Figures/test_cases.eps}
\caption{Two simulated images.
{\bf Left}: A field containing $\sim$100 stars.
{\bf Right}: A field containing $\sim$1000 stars.\label{fig:simulated_data}}
\end{center}
\end{figure}

@@

\section{Broken Power-Law Distribution}\label{power_law}
The broken power-law distribution is based on a straightforward extension
to a simple power-law distribution (also known as a Pareto distribution,
particularly in the statistics literature). The power-law distribution for a
variable $x$ (given a lower cutoff $x=h$ and a slope $\alpha$) is defined by:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcr}
0, & & x < h \\
x^{-\alpha - 1} & & x \geq h.
\end{array}
\right.
\end{eqnarray}
In contrast, the broken power-law distribution for a variable $x$ is defined by
a lower cutoff $x=h_1$, two slopes $\{\alpha_1, \alpha_2\}$ and a break point
$x=h_2$:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
x^{-\alpha_1 - 1} & & h_1 \leq x \leq h_2 \\
x^{-\alpha_2 - 1} & & x > h_2.
\end{array}
\right.
\end{eqnarray}
The free parameters of the broken power-law are:
\begin{eqnarray}
\beta = \{h_1, h_2, \alpha_1, \alpha_2\}.
\end{eqnarray}
With normalising terms included, the proportionality becomes an equality:
\begin{eqnarray}
p(x) &=&
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
Z_1^{-1}x^{-\alpha_1 - 1}, & & h_1 \leq x \leq h_2 \\
Z_2^{-1}x^{-\alpha_2 - 1}, & & x > h_2.
\end{array}
\right.
\end{eqnarray}
Two conditions will be used to determine the normalisers $Z_1$ and $Z_2$.
Firstly, the PDF should be continuous at $x=h_2$:
\begin{eqnarray}
Z_1^{-1}h_2^{-\alpha_1 - 1} &=& Z_2^{-1}h_2^{-\alpha_2 - 1}\\
\implies
Z_2 &=& Z_1h_2^{\alpha_1-\alpha_2}
\end{eqnarray}
The second condition is that the total probability must be 1:
\begin{eqnarray}
\int_{h_1}^{h_2} Z_1^{-1} x^{-\alpha_1 - 1} \, dx
+
\int_{h_2}^\infty Z_2^{-1} x^{-\alpha_2 - 1} \, dx
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
Z_2^{-1}\alpha_2^{-1}h_2^{-\alpha_2}
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
Z_1^{-1}h_2^{\alpha_2-\alpha_1}\alpha_2^{-1}h_2^{-\alpha_2}
&=& 1
\end{eqnarray}
\begin{eqnarray}
\implies
Z_1 &=& \alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
h_2^{-\alpha_1}\alpha_2^{-1}.
\end{eqnarray}
The cumulative distribution (CDF) is:
\begin{eqnarray}
P(X \leq x) = F(x) &=& 
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
(Z_1\alpha_1)^{-1}\left(h_1^{-\alpha_1} - x^{-\alpha_1}\right), & & h_1 \leq x \leq h_2 \\
1 - (Z_2\alpha_2)^{-1}x^{-\alpha_2}, & & x > h_2 
\end{array}
\right.
\end{eqnarray}
and the inverse CDF is:
\begin{eqnarray}
F^{-1}(u) &=& 
\left\{
\begin{array}{lcr}
\left[h_1^{-\alpha_1} - uZ_1\alpha_1\right]^{-1/\alpha_1}, & & 0 < u < 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}\\
\left[Z_2\alpha_2(1-u)\right]^{-1/\alpha_2},& & 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}< u < 1
\end{array}
\right.
\end{eqnarray}
An example of a broken power-law distribution is shown in Figure~\ref{fig:powerlaw}.
\begin{figure}[h!]
\begin{center}
\includegraphics[scale=0.5]{Figures/broken.eps}
\caption{A broken power-law distribution.\label{fig:powerlaw}}
\end{center}
\end{figure}

@@

\section{Usage}
Software can be obtained from {\tt http://github.com/dfm/star-field/}

\section{Proposal Distributions}

@@

\bibitem[Hogg and Lang(2011)]{2011EAS....45..351H} Hogg, D.~W., Lang, D.\ 
2011.\ Telescopes don't make catalogues!.\ EAS Publications Series 45, 
351-358.
